---
title: "Reg_Model_PA"
author: "wpshen"
date: "March 31, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary  
In this report, we explored the relationship between a set of variables and miles per gallon (MPG) (outcome) in the "mtcars"" dataset. The analysis provided the answers for the following two questions:

1. Is an automatic or manual transmission better for MPG?
2. Quantify the MPG difference between automatic and manual transmissions?

The results show that the manual transmissions does give cars a better MPG. The mean for MPG of manual, 24.39, is higher than that of automatic, 17.15.

This study also create the better fit linear model for the MPG data, using only the Weight(wt), Transmission(am), and 1/4 mile time(qsec) as the regressors (including the interaction between wt and am). The Adjusted R-squared value of the model is 0.8804, which means it can explain ~88% of the variance.

## Exploratory Data Analysis

From the help of mtcars, it is a data fram with 32 observations on 11 variables.

[, 1]	mpg	    Miles/(US) gallon
[, 2]	cyl	    Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	    Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	    Weight (1000 lbs)
[, 7]	qsec	1/4 mile time
[, 8]	vs	    V/S
[, 9]	am	    Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors

```{r}
data(mtcars)
head(mtcars)
```


## Is an automatic or manual transmission better for MPG?  
Using a quick boxplot to exam the effect of tranmission on MPG

```{r}
## am = Transmission (0 = automatic, 1 = manual)
dt <- subset (mtcars, select = c(mpg,am))
dt$am <- factor(dt$am, labels=c("Auto","Manual")) 
boxplot(mpg ~ am, data = dt, 
        xlab = "Transmission Type",
        ylab = "Miles per Gallon",
        main = "MPG by Transmission Type")
```

It seems clear the the Manual transmission gives the cars a better Miles per Gallon on average.

## Quantify the MPG difference between automatic and manual transmissions?  
To quantify the difference, we can perform a quick T-test. The null hypothesis is that the automatic or manual transmission has the same effect on the MPG.  

```{r}
t.test(mpg ~ am, data = dt)
```  

Since the p-value is 0.00137, the null hypothesis was rejected. So, the automatic transmissions does have an impact on the MPG. The mean for MPG of manual, 24.39, is higher than that of automatic, 17.15. Therefore, manual transmission is better for MPG.  


## Regression Modeling
Since we have decided to use linear model `lm` to fit the data, let's first do a variable selection. Two methods were used here have yield the same result. The first is to use the `varImp` function in the `caret` package and compute the importance of different variables against MPG.  

```{r}
library(mlbench)
library(caret)
control <- trainControl(method="repeatedcv", number=10, repeats=3)
model <- train(mpg~., data=mtcars, method="lm", preProcess="scale", trControl=control)
importance <- varImp(model, scale=FALSE)
plot (importance)
```

As shown, the Weight(wt), Transmission(am), and 1/4 mile time(qsec) are the most important variables. Alternatively, We can also use the stepAIC{MASS} fucntion for the analysis

```{r}
library(MASS)
step <- stepAIC(lm(mpg~.,data=mtcars), trace=FALSE)
step$anova
```

The anova result has recommended the use of Weight(wt), Transmission(am), and 1/4 mile time(qsec) to fit the linear model

```{r}
fit <- lm (mpg ~ wt + am + qsec, data = mtcars)
summary (fit)
```

We can further improve the model by looking at the interaction between variables. 

```{r}
cor (subset(mtcars, select=c("wt","am","qsec")))
```

It indicates that "wt" and "am" has an interaction, 0.69. So let's add this interaction and check the model again,

```{r}
betterfit <- lm (mpg ~ wt*am + qsec, data = mtcars)
summary (betterfit)
```
Now the Adjusted R-squared value is 0.8804, which means the model can explain ~88% of the variance, an improvement of ~5% from without the interaction. It is a pretty good model now.

```{r}
## Residual Plots
par(mfrow = c(2, 2))
plot(betterfit)
```  


