---
title: "Statistical Inference Project Part 2"
author: "wpshen"
date: "March 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This is the Part 2 of the project for the statistical inference class. We will 
1. Load the ToothGrowth data
2. Provide a basic summary of the data and perform some basic exploratory data analysis
3. Compare tooth growth by supp and dose
4. State conclusions and the assumptions needed to make them

## Exploratory Data Analysis

Load the ToothGrowth data and perform some basic exploratory data analyses.
```{r}
tr <- ToothGrowth
str(tr)
table(tr$dose)
```

A quick table() of the "does" shows that there are only three dose levels, i.e. 0.5, 1 and 2. Therefore, convert the numeric variable to a factor.

```{r}
tr$dose<-as.factor(tr$dose)
str(tr)
```

Here is a basic summary of the data and a boxplot for the Tooth Growth as a function of the suppliment and does.
```{r}
summary(tr)
library(ggplot2)
g <- ggplot(tr, aes(x=dose,y=len)) + facet_grid(.~supp) + geom_boxplot()
g + labs(title = "Boxplots by Supplement & Does", x = "Supplement & Dose",
         y = "Tooth Growth")

```

## Hypothesis test of the tooth growth by suppliment and dose

Using t.test, we will check if there's difference in the effect of gaving Orange Juicy and pure Vitamin C. The null hypothesis here is no difference and the p value threshold is set on 0.05. Any test that has the p value below the threshold will support us to reject the null hypothesis.

```{r}
  t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == .5, ]) 
  t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 1, ]) 
  t.test(len ~ supp, ToothGrowth[ToothGrowth$dose == 2, ]) 

```

##Assumptions and Conclusions

The dataset as presented in R has little supporting documentation, but we have assumed that the data is indeed paired.

From the exploratory data analysis, we observed that the tooth growth increases as the dosage of Orange Juice (OJ) or vitamin C (VC) increases.

For the T-test analysis above, our Null hypothesis is that there is no difference in growth impact by orange juice or Vitamin C. From the less statistically significant p-values (<0.05) that for dosages of 0.5 mg and 1 mg, we can reject the Null hypothesis. Meaning, the orange juice is more effective at promoting tooth growth than just Vitamin C. From the p-value for the 2 mg (0.964), we can conclude that orange juice and Vitamin C are equally effective when the dosage is 2 mg.
