---
title: "Statistical Inference Project Part 1"
author: "wpshen"
date: "March 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Overview
This is the Part 1 of the the statistical inference class project. In it, we will run simulations to investigate the exponential distribution in R and compare it with the Central Limit Theorem.. 

##Simulation
In R, he exponential distribution can be simulated with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. We will set the lambda = 0.2 for all of the simulations. The sample size n = 40 with number of simulation numsim = 1,000. 

In the following sections, we will show

1. The sample mean and compare it to the theoretical mean of the distribution.
2. how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. that the distribution is approximately normal.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

#### Initialize the variables and data set
```{r}
n<-40
lambda<-0.2
numsim<-1000
dataset<-matrix(rexp(n*numsim,lambda),numsim)
```

Plot the distribution of averages of 40 exponentials
```{r}
simMean <- data.frame(mean=apply(dataset,1,mean))
library(ggplot2)
g <- ggplot(simMean, aes(mean)) 
g <- g + geom_histogram(binwidth = lambda, color="white", aes(y = ..density..)) 
g <- g + labs(title="Distribution of averages of 40 exponentials", 
              x="Mean of 40 Exponentials",y="Density") 
g
```

### Compare the sample mean to the theoretical mean 
```{r}
theoryMean <- 1/lambda
sampleMean <- mean(simMean$mean)
cat("Theoretical Mean :",as.character(theoryMean),"\n",
    "Sample Mean :",strtrim(as.character(sampleMean),5), sep = "")

```


### Show the sample variance and compare it to the theoretical variance
```{r}
theoryVar<-(1/lambda)^2/n
sampleVar<-sd(simMean$mean)^2
cat("Theoretical Varience :",as.character(theoryVar),"\n",
    "Varience of Samples :",strtrim(as.character(sampleVar),5), sep = "")
```

### Exam whether the distribution is approximately normal
```{r}
g <- g + stat_function(fun = dnorm, args=list(mean=theoryMean, sd=sqrt(theoryVar)), colour = "red", size = 1.0)
g
```

As shown on the plot, the distribution is approximately normal 
